---
title: "Gov 50 Final Project"
author: "Shriya Yarlagadda"
description: "Gov 50 Final Project"
output:
  distill::distill_article:
    self_contained: false
---



## Milestone 1

## Project thoughts

Broadly, I'd like to study an internationally-focused topic. My interests are largely concentrated in two realms, which are human rights and sustainability, and I hope to study one of the two. Within human rights, I'd be interested in looking into United Nations voting patterns related to militarized conflict and their links to genocide or other human rights violations. Within sustainability, I'm interested in comparing sustainability policy across countries, perhaps related to developmental aid. I'm a bit more interested in human rights topics, but am excited to see which area I'd be able to find the most interesting data set in.

## Milestone 2

## Selecting Data
```{r Prepping for Data Loading}

# install.packages("readxl")
library("readxl")
library(tidyverse)

```

```{r Reading in Data}
svdata <- read_excel("SVAC_3.0_complete (1).xlsx", sheet = 1)
```

Data Citation: 
Cohen, Dara Kay and Ragnhild Nordås. 2014. Sexual Violence in Armed Conflict Dataset. [2022], from the Sexual Violence in Armed Conflict Dataset website: http://www.sexualviolencedata.org

Does the scale of a conflict impact the prevalence of sexual violence? Through my analysis, which uses the Sexual Violence Data Set, I hope to explore how the expansiveness of a violent conflict (that is, the breadth of the location of its actors) impacts the prevalence of acts of sexual violence as an act of war. My hypothesis is that the more expansive a conflict is (ie, if the actors are different states rather than actors within a state), the more likely it is that acts of sexual violence will be widespread. This is because I hypothesize that conflicts involving multiple states cover a wider geographical area, allowing for a higher prevalence of violent actions. My sample is recorded acts of sexual violence during armed conflict committed by armed actors between 1989 and 2019. My unit of analysis is armed conflicts between 1989 and 2019. My explanatory variable is the scale of the armed conflict during which an act of sexual violence is committed. This variable is coded 2 = Interstate Conflict, 3 = Intrastate Conflict, 4 = Internationalized Internal Armed Conflict. I intend to recode this variable so that it increases by the scale of the conflict (1 = Intrastate Conflict, 2 = Internationalized Internal Armed Conflict, 3 = Interstate conflict). My outcome variable is the expansiveness of sexual violence in an area. It is coded 0 = None, 1 = Isolated, 2 = Numerous, 3 = Massive, and -99 for no data. This variable is also disaggregated amongst three different sources and I will account for this by taking the highest number reported across the three sources (in line with the methodology of this data set's code, which codes for the higher number in this variable given conflicting information). I will see evidence that my hypothesis is correct if higher values of my explanatory variable (which suggests a wider scale of war) is associated with higher expansiveness of sexual violence (my outcome variable). If I see the opposite relationship, with higher values of conflict expansiveness being associated with lower rates of sexual violence, or, alternatively, no substantive relationship, then I will see that my hypothesis is incorrect. 

```{r Cleaning Data}
svdata <- svdata |>
  mutate(prev_agg = pmax(state_prev, ai_prev, hrw_prev, na.rm = TRUE)) |>
  filter(prev_agg != -99)

svdata <- svdata |>
  mutate(
    type_new = case_when(
      type == 3 ~ 1,
      type == 2 ~ 3,
      type == 4 ~ 2
    )
  )
  

svsubset <- svdata |>
  mutate(prev_agg = pmax(state_prev, ai_prev, hrw_prev, na.rm = TRUE)) |>
  filter(prev_agg != -99 & prev_agg != 0)



```

## Milestone 3

## Histogram of Violence - Dependent Variable

```{r}

ggplot(data = svdata, mapping = aes(x = prev_agg)) + 
  geom_histogram(binwidth = 1, color = "indianred1") + 
  labs(x = "Prevalence Rates of Sexual Violence in Armed Conflict" , y = "Count") +
  theme_bw()
```

## Milestone 4

## Scatterplots 

```{r}

sp1 <- svdata |> ##Year/Subset
  ggplot(aes (x = type_new, y = prev_agg)) +
  geom_point() +
  geom_jitter() +
  geom_smooth(method = "lm") + 
  labs(x = "Expansiveness of Conflict", y = "Prev. Sex. Violence") +
  theme_bw()

sp2 <- svsubset |> ## Expansiveness/Subset
  ggplot(aes (x = type_new, y = prev_agg)) +
  geom_point() +
  geom_jitter() +
  geom_smooth(method = "lm") + 
  labs(x = "Expansiveness of Conflict", y = "Prev. Sex. Violence (excluding 0s)") +
  theme_bw()

sp3 <- svdata |> ## Year/OG
  ggplot(aes (x = year, y = prev_agg)) +
  geom_point() +
  geom_jitter(width = 0.01) +
  geom_smooth(method = "lm") + 
  labs(x = "Year of Conflict", y = "Prev. Sex. Violence") +
  theme_bw()

sp4 <- svsubset |> ## Year/Subset
  ggplot(aes (x = year, y = prev_agg)) +
  geom_point() +
  geom_jitter(width = 0.01) +
  geom_smooth(method = "lm") + 
  labs(x = "Year of Conflict", y = "Prev. Sex. Violence") +
  theme_bw()


sp1

sp2

sp3

sp4
```

## Regression and Potential Confounders

```{r}

library(broom)
library(knitr)

reg1 <- lm(prev_agg ~ type_new, data = svdata)
reg2 <- lm(prev_agg ~ type_new, data = svsubset)

reg3 <- lm(prev_agg ~ type_new + year , data = svdata) 
reg4 <- lm(prev_agg ~ type_new + year , data = svsubset)

reg1|>
  tidy() |>
  kable(caption = "Coefficient Estimates for a Regression of Violence Rates and Conflict Expansiveness", 
        col.names = c("Predictor", "Estimate", "Std. Error", "Statistic", "P-value"),
        digits = c(0,2,3,2,3))

reg2|>
  tidy() |>
  kable(caption = "Coefficient Estimates for a Regression of Violence Rates and Conflict Expansiveness (Subset)", 
        col.names = c("Predictor", "Estimate", "Std. Error", "Statistic", "P-value"),
        digits = c(0,2,3,2,3))

reg3|>
  tidy() |>
  kable(caption = "Coefficient Estimates for a Regression of Violence Rates and Various Variables", 
        col.names = c("Predictor", "Estimate", "Std. Error", "Statistic", "P-value"),
        digits = c(0,2,3,2,3))

reg4|>
  tidy() |>
  kable(caption = "Coefficient Estimates for a Regression of Violence Rates and Various Variables (Subset)", 
        col.names = c("Predictor", "Estimate", "Std. Error", "Statistic", "P-value"),
        digits = c(0,2,3,2,3))
```

The first two regressions, which compare the full dataset (including conflicts in which sexual violence was not recorded as a form of conflict) and the subset, appear to present contradicting evidence. While the regression from the full dataset suggested that a one unit increase in the expansiveness of a conflict was associated with a 0.07 decrease in the prevalence of sexual violence, the regression from the subset suggested that a one unit increase in the expansiveness of a conflict was associated with a 0.13 increase in the prevalence of sexual violence. This appears to suggest that while, on aggregate, wider conflicts are associated with a slight decrease in sexual violence, this may be biased by wider conflicts being more likely to lack sexual violence at all. Thus, if you remove conflicts with no reports of sexual violence, there seems to be a positive relationship. However, this again appears to be misleading when reviewing the scatterplots, as there are far more smaller scale conflicts than larger ones. 

Furthermore, the scatterplot of year and prevalence of sexual violence appears to show an increasing trend, especially when using the subset data. When running a regression including this variable as well (ie holding year constant), there appears to only be a 0.09 increase in the prevalence of sexual violence given a more expansive conflict. Thus, this may be a confounding variable. 


## Final Project

## Introduction

Despite its illegality under the Geneva Convention and classification as a war crime, the use of sexual violence in armed conflict is unfortunately common (International Committee of the Red Cross, 2016). Given this commonality, it is incredibly useful to study its presence, as doing so may contribute to greater scholarly understanding of the phenomenon.Thus, through my work on this project, I hope to explore the research question "Does the expansiveness of a conflict impact the prevalence of sexual violence?" I hypothesize that the more expansive a conflict is, the more prevalent sexual violence will be. This is because I believe that the larger geographic area a conflict is waged upon, the more geographic area is available to commit acts of sexual violence. Furthermore, I hypothesize that individuals committing such acts may be less willing to commit acts of sexual violence in regions that are familiar to them – ie, those that are closer to their home location. Through better understanding the answer to this question, we can more effectively move towards useful future solutions.


## Data

To conduct this analysis, I used the Sexual Violence in Armed Conflict Dataset, which measures acts of sexual violence in "active state-based armed conflicts" between 1989 and 2019 (Cohen et al. 2021). While this dataset includes a number of other interesting variables, including actor types, conflict issue, and conflict years, I focus on analyzing the relationship between conflict type (my independent variable) and the prevalence of sexual conflict (my dependent variable). Both of these variables are derived through expert coding, using data from the UCDP/PRIO for the independent variable and three separate sources (US State Department annual reports, Amnesty International annual and special reports, and Human Rights Watch annual and special reports) for the dependent variable. Notably, these two selected variables benefit from being measured using data that is not closely related. In other words, the data that is used to classify a conflict's expansiveness is not the same data used to classify level of sexual violence. This limits confounding variables. 

For the context of this study, I slightly adjusted the provided coding. In order to represent an increase in conflict expansiveness, I recoded the independent variable such that a 1 represents an Intrastate Conflict, 2 represents an Internationalized Internal Armed Conflict, and 3 repesents an Interstate Conflict. In contrast to the previous coding (which instead used 2,3,and 4 and was ordered differently), this coding would better allow me to regress the relationship between an increase in size of conflict and other variables, as this order more closely predicts an increase in conflict area. While this measure may not be a perfect representation of the geographic size of a conflict, it is likely a closer approximation than the previous arrangement. 

Similarly, given the potential for variation amongst the three sources of information on the prevalence of sexual violence, I calculated an aggregate measure of the prevalence of sexual violence during my analysis that selected the highest score assigned by one of these sources, still using the same scale that was used in the dataset's codebook. In other words, if, for example, the US State Department and Amnesty International reports reported no sexual violence (coded as 0) but the Human Rights Watch reports reported some sexual violence (coded as 1), the aggregate measure would report a prevalence score of 1 for the associated conflict. Primarily, this allowed me to account for situations in which only one of the sources has information. However, I used the maximum rather than the average of the three sources given that data on sexual violence is often underreported (International Committee of the Red Cross, 2016). By using the highest of the available source, I hope to counteract this underrepresentation.

This research utilizes time-series cross sectional data (also known as Panel Data). However, for the context of this course, I am utilizing a cross-sectional research design. 

A plot summarizing the distribution of the dependent variable (the counts of the various levels of sexual violence prevalence) is found in the histogram below. It demonstrates that while most of the conflicts measured did not have recorded measures of sexual violence, there are an alarming number that do. While most acts are isolated if they occur, there are also a few incidents in larger scale. 

# Histogram of Prevalence of Sexual Violence
```{r}
ggplot(data = svdata, mapping = aes(x = prev_agg)) + 
  geom_histogram(binwidth = 1, color = "indianred1") + 
  labs(x = "Prevalence Rates of Sexual Violence in Armed Conflict" , y = "Count") +
  theme_bw()
```

## Results

The relationship between the expansiveness of conflict and the prevalence of sexual violence in conflict can be expressed through the following scatterplot. However, given the large number of conflicts with no violence, which may skew results, I also created a scatterplot with a subsetted version of the data, which excluded conflicts without any sexual violence. 

```{r}
sp1

sp2
```

Interestingly, while the original dataset suggests a positive relationship between expansiveness and prevalence of sexual violence, the subsetted dataset conversely suggests a negative relationship. This appears to suggest that while, on aggregate, wider conflicts are associated with less sexual violence, this may be biased by wider conflicts being more likely to lack sexual violence at all. Thus, if you remove conflicts with no reports of sexual violence, there seems to be a positive relationship. However, this appears to be biased by the fact that there are far more smaller scale conflicts than larger ones.

Given this visualization, I conducted regressions of these relationships to gain a greater understanding of their relationships. 


```{r}
reg1|>
  tidy() |>
  kable(caption = "Coefficient Estimates for a Regression of Violence Rates and Conflict Expansiveness", 
        col.names = c("Predictor", "Estimate", "Std. Error", "Statistic", "P-value"),
        digits = c(0,2,3,2,3))

reg2|>
  tidy() |>
  kable(caption = "Coefficient Estimates for a Regression of Violence Rates and Conflict Expansiveness (Subset)", 
        col.names = c("Predictor", "Estimate", "Std. Error", "Statistic", "P-value"),
        digits = c(0,2,3,2,3))
```

The regression of the original data suggests that a 1 unit increase in the expansiveness of a conflict results in a 0.07 decrease in the prevalence of sexual violence. However, the regression of the subsetted data suggests that a 1 unit increase in the expansiveness of a conflict results in a 0.13 increase in the prevalence of sexual violence. Given that both calculations have a p-value that is less than 0.05, we can reject the null hypothesis that there is no relationship between the two variables at an alpha level of 0.05. This implies that there is a positive relationship between the expansiveness of a conflict and the prevalence of sexual violence only when there is a record of acts of sexual violence. Otherwise, there relationship is broadly negative. 

Given this conclusion, I also recognized that there may be notable confounding variables. Thus, I began by creating scatterplots of the relationship between year and prevalence of violence. If this is not a confounding variable, there should be no relationship between the two. 

```{r}
sp3

sp4
```
In the case of the subset, there definitely appears to be a relationship between the variables from the scatterplot. However, it is not clear whether there is one in the original dataset. In order to better measure this, I calculated the following regressions, which included both year and expansiveness of conflict.

```{r}
reg3|>
  tidy() |>
  kable(caption = "Coefficient Estimates for a Regression of Violence Rates and Various Variables", 
        col.names = c("Predictor", "Estimate", "Std. Error", "Statistic", "P-value"),
        digits = c(0,2,3,2,3))

reg4|>
  tidy() |>
  kable(caption = "Coefficient Estimates for a Regression of Violence Rates and Various Variables (Subset)", 
        col.names = c("Predictor", "Estimate", "Std. Error", "Statistic", "P-value"),
        digits = c(0,2,3,2,3))
```
This suggests that in the original data, when the expansiveness of conflict is held constant, there is no relationship between year and prevalence of sexual violence. Thus, it suggests that, holding year constant, an increase in the expansiveness of conflict by one unit will result in a 0.08 decrease in the prevalence of sexual violence (slightly larger than in the original regression, due to rounding). In the subsetted data though, when expansiveness of conflict is held constant, a 1 unit increase in year results in a slight (0.02) increase in the prevalence of sexual violence. Similarly, when year is held constant, there is a 0.09 increase in the prevalence of sexual violence with every one unit increase in expansiveness of the conflict. Although this may point to the existence of confounding variable, especially in the latter case, both coefficient values for the variable "type_new" (the independent variable) have a p-value less than or equal to 0.05. Therefore, even though the value of the coefficient may have decreased, there still appears to be a significant relationship between expansiveness of a conflict and the prevalence of sexual violence. While we are not able to control for every potential confounding variable in this regression and are thus unable to claim that this relationship is causal, this work points to an interesting correlation. 

## Conclusion
Overall, there does not appear to be a conclusive relationship between the expansiveness of an armed conflict and the prevalence of sexual violence. Although there appears to be a slightly negative relationship between the two overall, there is a slightly positive relationship only considering only cases that did have incidents of sexual violence. While year could perhaps be considered a confounding variable, these relationships hold at an alpha level of 0.05 holding year constant, meaning that there is a relationship of note.

This analysis is limited by how each of the variables are defined and measured. Given that both are measured and coded by humans, there is a potential for numerical error. Beyond this, though, there may be inconsistencies in how the independent variable (the expansiveness of the conflict) are measured. While I chose to order the categories from intrastate, internationalized intrastate, and interstate, others may argue that this does not accurately represent "expansiveness." Furthermore, the expansiveness of a conflict may be changing throughout a conflict, and sexual violence rates may change accordingly.Perhaps then, a future study would be better served by additional data better quantifying geographic expansiveness of a conflict, such as the geographic area covered by battlefields. 


## Citations

International Committee of the Red Cross. (2016, August 19). *Q&A: Sexual violence in armed conflict*. https://www.icrc.org/en/document/sexual-violence-armed-conflict-questions-and-answers#:~:text=Rape%20and%20other%20forms%20of,the%20prohibition%20of%20sexual%20violence.

Cohen, Dara Kay, Ragnhild Nordås, and Robert Ulrich Nagel. 2021. *Sexual Violence in Armed Conflict Dataset 3.0*. [2022], from the Sexual Violence in Armed Conflict Dataset website: http://www.sexualviolencedata.org